Tidal streams in triaxial systems

### 2015-02-11 ms.tex
paper updates

All of these parameters are specified in terms of the \emph{density}; for orbit analysis, we need to determine the form of the potential in terms of these parameters, which, in general, requires numerical integration of the density at each position of interest. For computational efficiency, many authors instead express the triaxiality in the form of the potential, however this can lead to unphysical situations where the density becomes negative. \citet{lee03} derive a perturbative expansion of the potential integral and show that the expansion is accurate even for modest axis ratios (e.g., the median values shown above).

For this work, we use the triaxial potential expression from \citet{lee03}, parametrized in a slightly different manner. In terms of spherical coordinates\footnote{(radius, azimuth, colatitude)} with the radius normalized by the scale radius, $u = r/r_s$
\begin{align}
	\Phi(u,\phi,\theta) &\approx \frac{v_c^2}{A}\left[F_1(u) + \frac{1}{2}(e_b^2 + e_c^2)F_2(u) + \frac{1}{2} [(e_b\sin\theta \sin\phi)^2 + (e_c\cos\theta)^2] F_3(u) \right]\\
	A &= \left(\ln2 - \frac{1}{2}\right) + \left(\ln2-\frac{3}{4}\right) (e_b^2 + e_c^2)
\end{align}
where $v_c$ is the circular velocity at the scale radius, $r_s$, along the $x$-axis, $e_b = \sqrt{1 - (b/a)^2}$, and $e_c = \sqrt{1 - (c/a)^2}$. The functions $F_i(u)$ are given in the appendix of \cite{lee03}. We chose $r_s=20~{\rm kpc}$ and $v_c = 200~{\rm km}~{\rm s}^{-1}$ to [...]. Figure~\ref{fig:potential} shows the rotation curve of this potential along the three principal axes.

@@

We therefore take a simplified approach that does reproduce [...] N-body simulations for several test cases. For a given set of orbital initial conditions, we find the coordinates at the nearest pericenter, initialize a small ``ball'' of test particle orbits around this position, then follow the orbits of all of these test particles for some time. For a given mass scale, the ball particle position and velocity offsets are assumed to be Normal distributed away from the parent orbit with spreads set by the tidal radius in position and velocity scale in velocity. If $(\bs{x}_0,\bs{v}_0)$ are the parent orbit initial conditions at pericenter, then $(\delta\bs{x}_i,\delta\bs{v}_i)$ is the deviation vector of the $i$th particle and is drawn from
\begin{align}
	\delta\bs{x}_i &\sim \mathcal{N}(0, r_{\rm tide})\\
	\delta\bs{v}_i &\sim \mathcal{N}(0, \sigma_v)\\
	r_{\rm tide} &=|\bs{x}_0| \left(\frac{m}{3M_{\rm enc}(|\bs{x}_0|)}\right)^{1/3} \\
	\sigma_v &= |\bs{v}_0|\left(\frac{m}{3M_{\rm enc}(|\bs{x}_0|)}\right)^{1/3} 
\end{align}
where $M_{\rm enc}(r)$ is the mass enclosed of the host potential within radius $r$, and $m$ is the mass scale of the debris.

For this work, we take $m=10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass to estimate the above scales.

@@

@@

\acknowledgements
APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155. 
This research made use of Astropy, a community-developed core \texttt{Python} package for Astronomy \citep{astropy13}.
This work additionally relied on Columbia University's \emph{Hotfoot} and \emph{Yeti} compute clusters, and we acknowledge the Columbia HPC support staff for assistance. \\

>>>>

All of these parameters are specified in terms of the \emph{density}; for orbit analysis, we need to determine the form of the potential in terms of these parameters, which, in general, requires numerical integration of the density at each position of interest. For computational efficiency, many authors instead express the triaxiality in the form of the potential, however this can lead to unphysical situations where the density becomes negative. \citet{leesuto03} derive a perturbative expansion of the potential integral and show that the expansion is accurate even for modest axis ratios (e.g., the median values shown above).

For this work, we use the triaxial potential expression from \citet{leesuto03}, parametrized in a slightly different manner. In terms of spherical coordinates\footnote{(radius, azimuth, colatitude)} with the radius normalized by the scale radius, $u = r/r_s$
\begin{align}
	\Phi(u,\phi,\theta) &\approx \frac{v_c^2}{A}\left[F_1(u) + \frac{1}{2}(e_b^2 + e_c^2)F_2(u) + \frac{1}{2} [(e_b\sin\theta \sin\phi)^2 + (e_c\cos\theta)^2] F_3(u) \right]\\
	A &= \left(\ln2 - \frac{1}{2}\right) + \left(\ln2-\frac{3}{4}\right) (e_b^2 + e_c^2)
\end{align}
where $v_c$ is the circular velocity at the scale radius, $r_s$, along the $x$-axis, $e_b = \sqrt{1 - (b/a)^2}$, and $e_c = \sqrt{1 - (c/a)^2}$. The functions $F_i(u)$ are given in the appendix of \cite{leesuto03}. We chose $r_s=20~{\rm kpc}$ and $v_c = 200~{\rm km}~{\rm s}^{-1}$ to [...]. Figure~\ref{fig:potential} shows the rotation curve of this potential along the three principal axes.

@@

We therefore take a simplified approach that does reproduce [...] N-body simulations for several test cases. For a given set of orbital initial conditions, we find the coordinates at the nearest pericenter, initialize a small ``ball'' of test particle orbits around this position, then follow the orbits of all of these test particles for some time. For a given mass scale, the ball particle position and velocity offsets are assumed to be Normal distributed away from the parent orbit with spreads set by the tidal radius in position and velocity scale in velocity. If $(\bs{x}_0,\bs{v}_0)$ are the parent orbit initial conditions at pericenter, then $(\delta\bs{x}_i,\delta\bs{v}_i)$ is the deviation vector of the $i$th particle and is drawn from
\begin{align}
	\delta\bs{x}_i &\sim \mathcal{N}(0, r_{\rm tide})\\
	\delta\bs{v}_i &\sim \mathcal{N}(0, \sigma_v)\\
	r_{\rm tide} &=|\bs{x}_0| \left(\frac{m}{3M_{\rm enc}(|\bs{x}_0|)}\right)^{1/3} \\
	\sigma_v &= |\bs{v}_0|\left(\frac{m}{3M_{\rm enc}(|\bs{x}_0|)}\right)^{1/3} 
\end{align}
where $M_{\rm enc}(r)$ is the mass enclosed of the host potential within radius $r$, and $m$ is the mass scale of the debris. For this work, we take $m=10^4~\msun$ to represent globular-cluster-like progenitors, and use the spherically-averaged enclosed mass to estimate the above scales.

@@

We follow \citet[][hereafter MV96]{merritt96} and quantify the degree of mixing by computing the ``distance'' between the configuration space density of the ensemble of particles at a given time, $\rho(\bs{x},t)$, with that of a distribution that is uniform over the energy hypersurface, $\rho_{E_0}(\bs{x})$, where $E_0$ is the energy of the parent orbit and hypersurface of interest. Our method differs in a few ways from that of MV96. First, MV96 use isoenergy ensembles of particles for following the density because that is the only way to guarantee that, if chaotic, nearly all orbits in the ensemble will lie on chaotic orbits. We instead use ensembles with small ($\approx$1\%) spreads in energy, but we still expect that such an ensemble around a parent chaotic orbit will reach a fully mixed state. The final distribution, rather than fill the energy hypersurface, will instead have some thickness around this surface. Around a typical regular orbit, each individual orbit will fill its torus ergodically -- that is, a surface with lower dimension than the energy hypersurface -- and thus, a small enough ensemble will never get ``close'' to the fully mixed state as defined above because it will remain confined into a subspace on the energy surface. In practice, only strongly chaotic systems will approach $\rho_{E_0}(\bs{x})$; the rate will depend on how stochastic the parent orbit is (e.g., as measured by the Lyapunov time or frequency diffusion rate), and also on how ``trapped'' the chaotic region is (a small chaotic region wedged between stable tori will take longer to appear chaotic, even if internally the frequency diffusion rate is large). However the difference in configuration space density between an ensemble undergoing chaotic mixing versus an ensemble around a regular orbit may be detectable over a much shorter timescale. The fully-mixed, normalized, configuration-space density that fills a given energy hypersurface, $E_0$, is 
\begin{align}
	\rho_{E_0}(\bs{x}) &= \int \dd^3v \, \rho_0 \, \delta(H - E_0)\\
	&= \int \dd^3v \, \rho_0 \, \delta\left(|\bs{v}|^2/2 + \Phi(\bs{x}) - E_0\right)\\
	&= 4\pi \int \dd v \, v^2 \, \rho_0 \, \delta\left(v^2/2 + \Phi(\bs{x}) - E_0\right)\\
	&= A \sqrt{E_0 - \Phi(\bs{x})}
\end{align}
where $A$ is a constant determined such that $\int \dd^3\bs{x} \, \rho_{E_0}(\bs{x}) = 1$, and $\delta(\cdot)$ is the Dirac delta function.

Second, we compute the Kullback-Leibler divergence \citep[KL divergence;][]{kullback51} by estimating the configuration space density $\rho(\bs{x},t_j)$ using kernel density estimation (KDE) with the ensemble of particle positions at time $t_j$, then use Monte Carlo integration to evaluate the KL divergence rather than evaluate the integral over a three dimensional grid. The KL divergence, $D_{\rm KL}$, is a measure of the relative Shannon entropy between two normalized density functions, $p(x)$ and $q(x)$,
\begin{equation}
	D_{\rm KL} = \int \dd x \, p(x) \ln \frac{p(x)}{q(x)}
\end{equation}
We would like to compute 
\begin{align}
	D_{\rm KL}(t_j) &= \int \dd^3 \bs{x} \, \rho(\bs{x},t_j) \ln \frac{\rho(\bs{x},t_j)}{\rho_{E_0}(\bs{x})}
\end{align}
however, we only have an estimate of the density $\rho_{\rm KDE}(\bs{x},t_j) \approx \rho(\bs{x},t_j)$. Using samples from this KDE estimated density, $\bs{x}_i$, the KL divergence is
\begin{align}
	D_{\rm KL}(t_j) &\approx \frac{1}{N} \sum_i \ln \frac{\rho_{\rm KDE}(\bs{x}_i,t_j)}{\rho_{E_0}(\bs{x}_i)}\\
	\bs{x}_i &\sim \rho_{\rm KDE}(\bs{x},t_j)\label{eq:sampled_from}
\end{align}
where Equation~\ref{eq:sampled_from} indicates that the $\bs{x}_i$ are samples from the KDE density. For any computation of the density estimate, we use 10-fold cross-validation to find the optimal kernel bandwidth, $\sigma_{\rm KDE}$, by maximizing the likelihood of withheld particles. Figure~\ref{fig:DKL} shows $D_{\rm KL}$ as a function of time for a regular (black) and chaotic (blue) orbit. The chaotic orbit ensemble mixes much faster than the regular orbit ensemble.

\begin{figure*}[!h]
\begin{center}
\includegraphics[width=0.75\textwidth]{figures/D_KL.pdf}
\caption{This is really just a placeholder that shows what I want to show here...} \label{fig:DKL}
\end{center}
\end{figure*}

@@

\acknowledgements
The authors wish to thank ...
APW is supported by a National Science Foundation Graduate Research Fellowship under Grant No.\ 11-44155. 
This research made use of Astropy, a community-developed core \texttt{Python} package for Astronomy \citep{astropy13}.
This work additionally relied on Columbia University's \emph{Hotfoot} and \emph{Yeti} compute clusters, and we acknowledge the Columbia HPC support staff for assistance.
